The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-07-06 manuscript.Rmd
edit manuscript

There can be public involvement in species description, but the exploration and collecting may remain an affair of specialists as it requires manipulating technical gear of have expertise. There are still many species that are easy to access to discover, but the bigger gap is in the species that are difficult to access, and small.

-->

@@

Furthermore, there are important discrepancies in our level of knowledge across phyla. While larger, and accessible groups represent the bulk of the records for marine invertebrates in museum collections, the small organisms, including representatives of the meiofauna are virtually absent. Of the `r n_idigbio_phyla` phyla occurring in iDigBio, `r records_rare_phyla$n_rare_phyla` are represented by less than 500 records, and make up `r records_rare_phyla$prop_rare_phyla`% of the total number of records.

<!-- From Mark: These previously undersampled taxa are likely to be much more important for the detection of subtle changes in environmental quality due to their size and physiological vulnerability. (?).  -->
<!-- also, the impact of changes in the composition in their communties is largely unexplored, look into Bik's paper? -->

<!-- - get numbers of species described by year, look if there is a correlation between sample numbers and number of new species (possible lag with time for description) -->
<!-- - see if proportion of species identified at genus (or higher) only has increased recently -->

@@

@@

iDigBio:
OBIS:
WoRMS:

@@

>>>>

@@

Furthermore, there are important discrepancies in our level of knowledge across phyla. While larger, and accessible groups represent the bulk of the records for marine invertebrates in museum collections, the small organisms, including representatives of the meiofauna are virtually absent. Of the `r n_idigbio_phyla` phyla occurring in iDigBio, `r records_rare_phyla$n_rare_phyla` are represented by less than 500 records, and make up `r records_rare_phyla$prop_rare_phyla`% of the total number of records. Yet, these undersampled taxa are likely to play an important role in the detection of changes in environmental quality due to their size, and physiological vulnerability [@Bik2012]. Additionally, the impact of these changes on higher trophic levels and ecosystems is largely unexplored.

To assess the status of taxonomic expertise in the US, we calculated the proportion of specimens identified at the species level through time. For all taxa, excepts for molluscs, there is a consistent decline, indicating that specimens remain for the most part unidentified in collections \@ref{fig:identification-through-time}. For diverse groups such as annelids or arthropods, the proportion of specimens identified at the species-level is now below 30%. The number of scientists describing new species might be increasing globally [@Costello2013], but local experts are needed to keep identifying the species that get collected.

```{r id_though_time, fig.cap='Proportion of marine invertebrate specimens collected through time identified at the species level. Trend lines are  B-splines of the 3rd degree polynomial. For all groups, we observe a decline in the proportion of specimens collected identified at the species level suggesting a lack of expertise to identify precisely what is collected. (\\#fig:identification-through-time)'}
knitr::include_graphics("figures/idigbio_identification_level_through_time.pdf")
```

@@

Crowd-sourced data and citizen science initiatives have been proposed to augment data about species abundance and distributions [@Pimm2014]. For marine taxa, beyond the application to near-shore habitats on easily recognizable species [@Delanay2008], large-scale efforts are currently not possible given the specialized equipment needed to sample most marine habitats and the expertise required to identify species. There however opportunities to utilize citizen science to improve biodiversity data and engage the general public into marine conservation issues. For instance it has been successfully used to classify plankton images [@Robinson2017], transcribe hand-written museum labels [@Hill2012], or be involved in bioblitzes [@Ballard2017]. These initiatives however involve a lot less people compared to projects related to birds.

@@

@@

<div id="refs"></div>

\newpage

## Supplementary material
